---
title: "TK8117_Week11_Group4: GLOBAL SENSITIVITY ANALYSIS"
author: "N.Quattrini"
date: "2024-11-05"
output: pdf_document
---

SOBOL'S GLOBAL SENSITIVITY ANALYSIS using airquality dataset
- contains daily air quality measurements in New York from May to Semptember 1973
- Q: how do different factors affect the Ozone concentration? 

# PACKAGES
```{r calling packages}
library(sensitivity)
library(boot)
library(ggplot2)
```

# STEP 1: LOADING & PREPARING DATA
  - importing dataset: 153 rows, 6 cols
  - removing rows containing NAs (42) --> 111 rows, 6 cols remain
```{r}
data(airquality)
airquality <- na.omit(airquality) 
```

# STEP 2 DEFINE THE MODEL
Creating simple linear model to predict Ozone concentration based on Solar.R, 
  Wind, and Temp
```{r model}
model <- function(X){
  fit <- lm(Ozone ~ Solar.R + Wind + Temp, data = airquality)
  predict(fit, newdata = data.frame(Solar.R = X[,1], Wind = X[,2], 
                                    Temp = X[,3]))
}
```

# STEP 3: GENERATE SAMPLES
```{r}
n <- 1000
X1 <- data.frame(matrix(runif(3 * n, 
                              min = apply(airquality[, c("Solar.R", 
                                                         "Wind", 
                                                         "Temp")], 2, min), 
                              max = apply(airquality[, c("Solar.R", 
                                                         "Wind", 
                                                         "Temp")], 2, max)), 
                        nrow = n))
X2 <- data.frame(matrix(runif(3 * n, 
                              min = apply(airquality[, c("Solar.R", 
                                                         "Wind", 
                                                         "Temp")], 2, min), 
                              max = apply(airquality[, c("Solar.R", 
                                                         "Wind", 
                                                         "Temp")], 2, max)), 
                        nrow = n))
```

# STEP 4: PERFORMING SENSITIVITY ANALYSIS
```{r sobol_result}
sobol_res <- sobol(model = model, X1 = X1, X2 = X2, order = 1, nboot = 100)
print(sobol_res)

# Example data from Sobol analysis
sobol_result <- list(S1 = c(0.3, 0.4, 0.2), ST = c(0.5, 0.6, 0.3), 
                     parameters = c("Solar.R", "Wind", "Temp"))
print(sobol_result)
```

# STEP 5: VISUALISING RESULTS
```{r plot}
# Convert the results to a data frame
df <- data.frame(
  Parameter = rep(sobol_result$parameters, 2),
  Sensitivity = c(sobol_result$S1, sobol_result$ST),
  Type = rep(c("First-order", "Total effect"), 
             each = length(sobol_result$parameters))
)

# Create the plot
g <- ggplot(df, aes(x = Parameter, y = Sensitivity, fill = Type)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Sobol Sensitivity Analysis", 
       x = "Parameters", y = "Sensitivity Index") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1")

ggsave("/Users/nicoleq/Library/CloudStorage/OneDrive-NTNU/LUX_NQJ/Courses/TK8117/sobol.jpg", plot = g, width = 6, height = 4, dpi = 300)


```



















